The low dimensional world of echolocating bats

### 2020-03-23 Paper/main_hp.tex
Herberts edits

One mechanism for place recognition would be for bats to parse the stream of echoes returning from the environment into localized and identified objects \citep{Lee2017,Barchi2013,Moss2001,Schnitzler2003,Simmons2012,Ulanovsky2008,Clare2015,Surlykke2016,Geipel2013}. However, as we argued before, it is highly questionable whether bats can interpret echoes in terms of a 3D environment model or acoustic image \citep[e.g.,][]{Vanderelst2015,Vanderelst2016,Steckel2013}. Sonar maps the three-dimensional structure of the environment on two one-dimensional waveforms arriving at the ears. While the time of arrival of echoes encodes object distance, azimuth and elevation directions are not directly available from the acoustic signal. In principle, azimuth and elevation can be estimated using spectral cues imposed by the head-related transfer function \citep[e.g.,][]{Reijniers2010,Aytekin2004}. However, complex environments, and even single plants \citep{Yovel2008,Muller2000}, return many overlapping and interfering echoes, making reconstructing the direction of origin of echoes much harder, and perhaps impossible. Extracting spatial information is further hampered by the limited temporal resolution of the sonar system \citep{Simmons1989,Wiegrebe1996,Surlykke1996}.

@@

Low dimensional descriptions potentially contain a lot of information \citep{Kuc1997b,Kuc1997}. We previously demonstrated that bats might use sonar template-based scene recognition \citep{Vanderelst2016,Vanderelst2017}. In this approach, bats are assumed to compare the acoustic signature of the environment to a set of previously stored signatures. We operationalized these acoustic signatures as time-intensity profiles, which could be derived directly from the cochlear output.

@@

Together these results suggest that echolocating bats live in a low dimensional perceptual world. The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of independent features, irrespective of whether we use artificial or real echo sequences. Nevertheless, this limited set of features is sufficiently rich for bats to recognize locations, and as we discuss, support flight control.

@@

Each set of points was converted to an impulse response with a sampling frequency $f_s$ of 219 kHz. The impulse response length was set to 7499 samples. These parameters correspond to the recording length and the sampling frequency used in the acoustic data collection (see below). In converting the point to an impulse response, we modeled attenuation of the echoes due to spherical spreading (both ways). We sampled the amplitude of the impulses from a normal distribution with a standard deviation of 10 dB.

The impulse responses were converted to echo sequences by convolving them with a bat-like emission. We used a 1 millisecond long linear frequency-swept cosine starting at 100 kHz and ending at 40 Khz. Again, these parameters correspond to those used in the acoustic data collection described below. The echo sequences were dechirped as follows, 
%
%ifft(complex conj(ftt emission) .* fft(echo))
%
\begin{equation}\label{key}
E_d \mathcal{F}^{-1}(\mathcal{F}(S)^* . \mathcal{F}(E))
\end{equation}
with $\mathcal{F}$ denoting the Fourier transform and $()^*$ the complex conjugate. The symbols $S$, $E$, $E_d$ refer to the emission, the echo sequence and the dechirped echo sequence, respectively. Examples of the resulting echo sequences are shown in figures \ref{fig:processing} and \ref{fig:synthetictemplates}.

Artificial echo sequences were filtered using a model of the bat's auditory periphery, as proposed by \citet{Wiegrebe2008}. This model simulates the temporal and spectral resolution of the bat's cochlea. The model consists of a gammatone filterbank followed by half-wave rectification and non-linear compression with an exponent of 0.4. Finally, the model includes a 1 kHz low-pass filter modeling the response characteristics of the inner hair cells. In the current implementation, we used a gammatone filterbank with xx filters, centered around xx kHz. The filters were spaced to correspond equivalent rectangular bandwidths \citep{Glasberg1990}. The output of each channel was filtered with a time (distance) weighted factor corresponding to the atmospheric attenuation at the channel's center frequency. This filtering approximated frequency-dependent atmospheric attenuation. The first 6 ms of data were removed. This was done to ensure the data corresponded with the acoustic data. As indicated below, we removed the first 6 ms of those data to avoid including the pickup of the emitted signal.

Figure \ref{fig:processing} depicts a single example of a synthetic echo sequence. It also illustrates the resulting response for each of the xx channels of the \citet{Wiegrebe2008} model. From these responses, it can be seen that these are highly correlated across channels. This is confirmed when analyzing the average cross-channel correlation for all xx synthetic echo sequences (Fig. \ref{fig:processing}xx). From these high cross-channel correlations, we conclude that the output of the cochlear model can be summarized by averaging across channels without extensive loss of information. On average, xx\% of variation in the output of the frequency channels was explained by the average across channels. An example of such a cross-channel sum is plotted in figure \ref{fig:processing}xx. In correspondence to our previous work \citep{Vanderelst2016}, we refer to this average across frequency channels as a template.

@@

We used \PCA\ to assess the dimensionality of the templates generated from the synthetic echo sequences. We found that almost all of the variance ($> 99\%$) in the templates could be captured by \pca\ components (see figure \ref{fig:processing}xx). Some examples of templates reconstructed from the principal components are shown in figure \ref{fig:synthetictemplates}xx).

\begin{figure*}[tb]
	\centering
	\includegraphics[width=1\linewidth]{../results/synthetic_templates}
	\caption{Example of synthetic dechirped echo sequences, their corresponding templates and the templates reconstructed from the \pca\ \pcs. (xx-xx) Three examples of synthetic echo sequence waveforms. Panels xx-xx depict the resulting templates (i.e., summation across the output channels of the \citet{Wiegrebe2008} model). These panels also show the result of projecting the templates into a \pca D space and transforming the result back to the original coordinates.}
	\label{fig:synthetictemplates}
\end{figure*}

In this section, we  used a generative model for naturalistic echoes. We assumed that echoes sequences received by bats can be considered as generated by anisotropically distributed point reflectors. Using this model, we have derived a low dimensional space ($D = \pca$), which can be used to describe the templates resulting from the artificial echo sequences. This implies that, after filtering through (a model of) the bat's auditory system, these artificial echoes can be converted to low dimensional descriptions. The original echo sequences, generated by xx to xx point reflectors effectively map to \pca\ dimensional vector. In the remainder of the paper, we will assess whether this low dimensional space allows describing the echoes from real bat habitats as collected using an ensonification device.

@@

In our previous work, the ensonification device was used to ensonify bat habitats at three different sites. However, here we only use data collected along two linear pathways through natural habitats. At the first site, a public park in Sde Boker, Israel (Fig. \ref{fig:datacollection}). The environment consisted of a walkway lined with rocks and vegetation. Unpublished bat monitoring data at the site confirmed that bat used this corridor to commute. The ensonification device was placed at 50 positions along a straight line spanning a total of 10 meters Locations were spaced 20 cm apart. At each location, echoes were collected for 31 azimuth directions (-150 to +150 degrees) and 7 elevation directions (-40 to +25 degrees). Therefore, at each location, echoes were collected from 31 x 7 = 217 directions. At each azimuth and elevation direction, three repeat measurements were recorded. In this way, a total of 32,550 echo trains (50 positions x 217 directions x 3 repeats) for each of the 31 microphones in the device were collected.

@@

As we did for the synthetic echoes sequences, the echoes trains collected in bat habitats were dechirped and filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}. We concluded from analyzing the synthetic echoes that the response in different frequency channels is highly correlated. Therefore, the output of the cochlear model was averaged across frequencies to obtain a time-intensity profile, or template. The beam of the emitter used was more narrow than the typically combined hearing and emission directionality of bats \citep{Vanderelst2010a,Reijniers2010,Jakobsen2012}. Therefore, templates were averaged across three neighboring directions. The first 6 ms of data were removed to avoid using emitted signals for analysis. The processing steps outlined above resulted in 3 $\times$ 217 templates, for each location at each site, corresponding to 217 directions and three repeats.

Among the sets of templates, some had very little energy in them, notably those collected for directions without reflectors. For example, aiming the ensonification device upwards often resulted in no echoes returning. These templates could not be classified \citep{Vanderelst2016}. Therefore, we removed them from the current analyses by omitting all templates with very low energy. The resulting number of templates included in the present analyses are given in table \ref{tab:sizes}.

@@

Internal (and, potentially, external) noise results in template values greater than zero, even in the absences of reflectors returning echoes. To avoid training the neural network on this noise, we estimated the noise floor in our data by collecting templates while pointing the device to empty space. The maximum template value obtained from these template measurements was used as the noise floor $\noisefloor$. The value of the noise floor $\noisefloor$ used had a numerical value of 0.15. Any template values below this value were set to this value. As we argued before \citep{Vanderelst2016}, using the empirical measurement noise to estimate $\noisefloor$ is biologically warranted as bats have a better signal to noise ratio in their auditory systems compared to our measuring device. Indeed, their dynamic range is at least 30 \db\ larger than our measurement device.

@@

We assessed whether the templates projected into the \pca-dimensional space still contained sufficient information to be recognized. Previously, we assessed whether templates can be used to recognize locations using, what \citet{Baddeley2012} called, a perfect memory approach. Templates were stored in a lookup table. The Euclidean distance between two echo templates was the metric used to assess how well bats should be able to distinguish between templates. This table lookup and distance calculation require a comprehensive search of memory. Hence, in our previous work, we established that templates could be classified, but this required a significant amount of memory and computational resources.

Here, we used a more biologically plausible memory system: an artificial neural network. These are a neurally inspired way of distributed information storage in artificial neural networks \citep{Mcleod1998}. However, we acknowledge that current training methods most likely have no biological counterpart.

Simple artificial feedforward neural networks were implemented using Tensorflow 2 for Python. The first layer of the neural network added Gaussian noise to each input sample ($\sigma(\stochasticnoise) = 0.01$). The following layers were completely connected and used the rectified linear activation function. The output layer used a softmax activation function (or the normalized exponential function), normalizing the output into a probability distribution consisting of probabilities proportional to the exponentials of the output layer's activation. The loss was calculated as the cross-entropy between the target output and the actual output of the neural network. Weights were updated after each epoch using the Adam algorithm \citep{Kingma2014} .

@@

Training a different neural network for each of these features ensured that the training algorithm paid equal attention to each feature. When training a single network for all features, the optimizer could preferably reduce errors in the location predictions as this feature has the largest variance. The full architecture can be considered as a single network with different streams for each feature (fig. \ref{fig:networks}).

@@

We found that the templates derived from acoustic data collected in real habitats could be well represented by the \pca\ components derived from synthetic data. This \pca-dimensional space could represent xx\% of the variance in the templates. Figure \ref{fig:empiricaltemplates} shows examples of templates and their reconstructed counterparts. Therefore, although the model used to generate echo sequences was very simple, the resulting underlying \pca D feature space could adequately represent templates resulting from real acoustic data.

@@

In our previous work \citep{Vanderelst2016}, we showed that the templates could be used to recognize locations in bat habitats. As we have shown that the templates can be projected into a lower \pca D feature space without loss, it can be expected that these \pca\ features can be used to locations as well. We found this to be indeed the case.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\linewidth]{../results/royal_results.pdf}
	\caption{Simular as fig \ref{fig:israelperformance}, but for the Royal Fort Park site.}
	\label{fig:royalperformance}
\end{figure*}

@@

\citet{Reijniers2010a} quantified the transmission of auditory information in the cochlea of bats. In that study, the authors adapted the Meddis inner hair cell model \citep{Meddis2006} to assess the amount of information in the acoustic signal preserved by the cochlea. These authors concluded that spectral notches introduced by overlapping echoes are faithfully encoded by auditory nerve spike patterns. However, quantifying the information transfer rate, they also found that about half of the information (in terms of bits) are lost in the transduction from acoustic energy to spike trains. In the current paper, we expand on this work and quantify the dimension reduction due to cochlear encoding. Moreover, we show that this reduction in complexity still supports an ecologically valid task: sonar-based scene recognition.

We assumed echoes received by bats can be modeled as originating from a large collection of point reflectors. Passing these synthetic echoes through a model of the auditory periphery, we obtained a time-intensity profile (template) for each echo sequence. These templates could be well described by a low dimensional space of \pca\ principal components (figure \ref{fig:synthetictemplates}). We found that the same \pca D space could be used to describe templates originating from echoes collected in real bat habitats (figure \ref{fig:empiricaltemplates}). This suggests the filtering of the auditory periphery provides the bat with a greatly reduced acoustic description of its environment. To identify the locus of the compression, we performed a number of additional analyses. In particular, we evaluated whether the low-dimensional description is a property of the echo sequences, the gammatone filters, or the 1 kHz low-pass filter in the \citet{Wiegrebe2008} model.

@@

Figure \ref{fig:mechanism} shows the number of \pcs\ needed to capture 99\% of the templates variance. From this plot, it can be seen that the number of \pcs\ increases when shuffling the echo sequences from xx to xx. Omitting the 1 kHz filter from the  \citet{Wiegrebe2008} model increases the number of \pcs\ to xx. Inspecting the average spectrum of the templates (fig. \ref{fig:mechanism}, inset) shows that most energy in the templates is contained in frequencies below 1 kHz. Therefore, if we take 1 kHz as the highest relevant frequency, it ought to be possible to completely represent templates by sampling them at 2 kHz. The corresponds to about 70 samples per template, as templates consist of 7499 samples at 219 kHz sample rate. This corresponds roughly to the number of \pcs\ required to represent templates derived from echo sequences that were shuffled  (fig. \ref{fig:mechanism}). The number of \pcs\ required to represent unshuffled templates is about half of this, i.c., \pca. Therefore, we conclude that the 1 kHz filter is the main, but not exclusive, locus of the dimensionality reduction. The structure of the echo signals also decreases the dimensionality of the templates. Therefore, we conclude that the sonar system of bats, through a number of filtering stages, compresses the 3D complexity of the world onto a lower dimensional representation.

@@

Despite the low dimensional character of the templates, they are sufficient to recognize environments. Indeed, this was demonstrated in our previous work \citep{Vanderelst2016}. Here, we directly demonstrated that the \pca\ principal components are sufficient to discriminate their associated locations using neural networks. The neural networks' performance was somewhat that the perfect memory approach we used previously (figs. \ref{fig:israelperformance}, \ref{fig:royalperformance}).

We showed that a low dimensional signal is sufficient for recognizing environments. This is a critical component of navigation. However, bats do more than navigating. They also rely on sonar for flight control. This includes, avoiding obstacles or following guiding structures. The environmental description the bat has access to for flight control has the same low-level character than that used for place recognition. Therefore, our current work suggests that this low-level description should be sufficient for flight control as well. Indeed, we have previously demonstrated that a low-dimensional description could be sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}. In that work, we converted both the echoes at the left and at the right ear to template-like representation. Subsequent work will have to demonstrate that the parameters of the \pca D space presented here can be used directly for controlling flight. In other words, we conjecture it should be possible to extract \pca\ features from an echo signal and convert this vector into a motor command. One can presume a neural network that converts the \pca\ features to flight control parameters.

@@

We were unable to find numbers on the number of neurons in the cortex of bat species. However, the mouse cortex has been estimated to contain about $9 \cdot 10^4$ neurons or $7 \cdot 10^8$ synapses per cubic millimeter \citep{Braitenberg2013}. The neural networks presented in this paper consisted of about \nneurons\ neurons and \nweights\ connections. Therefore, in as much as these neural networks can be considered as biologically plausible models of memory, they are orders of magnitude smaller than a single cubic millimeter of mouse cortex. Let alone the complete cortex, which has been estimated to contain 75 million neurons \citep{Williams2000}.

@@

It has been suggested before that bats' sonar systems provides detailed information about the world. A number of authors have referred to sonar as providing the bat with an acoustic image \citep{Lee2017,Barchi2013,Moss2001,Schnitzler2003,Simmons2012,Ulanovsky2008,Clare2015,Surlykke2016,Geipel2013}. As mentioned above, this is at odds with some recent behavioral and neurophysiological findings in bats faced with complex echo sequences \citep{Knowles2015,Geberl2013,Warnecke2018}. However, it is also at odds with the various filtering stages involved in the emission and the reception of bio-sonar. In particular, we argued before that the neurophysiology of the  bats auditory system represents a bottleneck in the amount of information accessible to the bat. Indeed, \citet{Reijniers2010a} already noticed that significant amounts of information get lost during cochlear transduction. Here, we followed up on that work to quantify the dimensional reduction of the input signals as a result of auditory processing.

From our current work, in combination with previous results \citep{Vanderelst2016,Vanderelst2015a,Mansour2019,Reijniers2010a}, we conclude that the sonar system of bats maps complex echoes to a limited number of features. These features can be efficiently stored to recognize scenes. Moreover, we conjecture that a (binaural) version of the features should allow for sonar-based flight control. As such, a view of echolocating bats living in a low dimensional sensory space emerges that contradicts existing claims of sonar providing acoustic images.

>>>>

One mechanism for place recognition would be for bats to parse the stream of echoes returning from the environment into localized and identified objects \citep{Lee2017,Barchi2013,Moss2001,Schnitzler2003,Simmons2012,Ulanovsky2008,Clare2015,Surlykke2016,Geipel2013}. However, as we argued before, it is highly questionable whether bats can interpret echoes in terms of a 3D environment model or acoustic image \citep[e.g.,][]{Vanderelst2015,Vanderelst2016,Steckel2013}. Sonar projects the three-dimensional structure of the environment on two one-dimensional waveforms arriving at the ears. While the time of arrival of echoes encodes object distance, azimuth and elevation directions are not directly available from the acoustic signal. In principle, azimuth and elevation can be estimated using spectral cues imposed by the head-related transfer function \citep[e.g.,][]{Reijniers2010,Aytekin2004}. However, complex environments, and even single plants \citep{Yovel2008,Muller2000}, return many overlapping and interfering echoes, making reconstructing the direction of origin of echoes much harder, and perhaps impossible. Extracting spatial information is further hampered by the limited temporal resolution of the sonar system \citep{Simmons1989,Wiegrebe1996,Surlykke1996}.

@@

However, it should be noted that such low dimensional descriptions have been shown to potentially contain a lot of useful information \citep{Kuc1997b,Kuc1997}. We previously demonstrated that bats might use sonar template-based scene recognition \citep{Vanderelst2016,Vanderelst2017}. In this approach, bats are assumed to compare the acoustic signature of the environment to a set of previously stored signatures. We operationalized these acoustic signatures as time-intensity profiles, which could be derived directly from the cochlear output.

@@

Together these results suggest that echolocating bats live in a low dimensional perceptual world. The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of independent features, irrespective of whether we use artificial or real echo sequences. Nevertheless, this limited set of features is sufficiently rich for bats to recognize locations, and as we discuss, could also support flight control.

@@

Each set of points was converted into an environmental impulse response with a sampling frequency $f_s$ of 219 kHz. The environmental impulse response length was set to 7499 samples. These parameters correspond to the recording length and the sampling frequency used in the acoustic data collection (see below). In representing each point reflector by a Dirac impulse, we modeled attenuation of the echoes by introducing an amplitude calculated based on spherical spreading (both ways). The final amplitude of each Dirac impulse was determined by sampling from a normal distribution centered on the calculated amplitude and with a standard deviation of 10 dB. This final operation represents a very simple model of varying reflector strength due e.g. to shape and orientation of the reflecting feature.

The impulse responses were converted to echo sequences by convolving them with a bat-like emission. We used a 1 millisecond long linear frequency-swept cosine starting at 100 kHz and ending at 40 Khz. Again, these parameters correspond to those used in the acoustic data collection described below. The echo sequences were dechirped as follows, 
%
%ifft(complex conj(ftt emission) .* fft(echo))
%
\begin{equation}\label{key}
E_d = \mathcal{F}^{-1}((\mathcal{F}(S))^* . \ \mathcal{F}(E))
\end{equation}
with $\mathcal{F}$ denoting the Fourier transform and $()^*$ the complex conjugate. The symbols $S$, $E$, $E_d$ refer to the emission, the echo sequence and the dechirped echo sequence, respectively. Examples of the resulting echo sequences are shown in figures \ref{fig:processing} and \ref{fig:synthetictemplates}.

Artificial echo sequences were filtered using a model of the bat's auditory periphery, as proposed by \citet{Wiegrebe2008}. This model simulates the temporal and spectral resolution of the bat's cochlea. The model consists of a gammatone filterbank followed by half-wave rectification and non-linear compression with an exponent of 0.4. Finally, the model includes a 1 kHz low-pass filter modeling the response characteristics of the inner hair cells. In the current implementation, we used a gammatone filterbank with xx filters, centered around xx kHz. The filters were spaced to correspond to equivalent rectangular bandwidths \citep{Glasberg1990}. The output of each channel was attenuated with a time (distance) weighted factor corresponding to the atmospheric attenuation at the channel's center frequency. This operation approximated frequency-dependent atmospheric attenuation. The first 6 ms of data were removed. This was done to ensure the data corresponded with the measurement data. As indicated below, we removed the first 6 ms of the measurement  data to avoid including the pickup of the emitted signal.

Figure \ref{fig:processing} depicts a single example of a synthetic echo sequence. It also illustrates the resulting response for each of the xx channels of the \citet{Wiegrebe2008} model. From these responses, it can be seen that they are highly correlated across channels. This is confirmed when analyzing the average cross-channel correlation for all xx synthetic echo sequences (Fig. \ref{fig:processing}xx). From these high cross-channel correlations, we conclude that the output of the cochlear model can be summarized by averaging across channels without extensive loss of information. On average, xx\% of variation in the output of the frequency channels was explained by the average across channels. An example of such a cross-channel sum is plotted in figure \ref{fig:processing}xx. In correspondence to our previous work \citep{Vanderelst2016}, we refer to this average across frequency channels as a template.

@@

We used principal components analysis (\PCA\ ) to assess the dimensionality of the templates generated from the synthetic echo sequences. We found that almost all of the variance ($> 99\%$) in the templates could be captured by \pca\ components (see figure \ref{fig:processing}xx). Some examples of templates reconstructed from the principal components are shown in figure \ref{fig:synthetictemplates}xx).

\begin{figure*}[tb]
	\centering
	\includegraphics[width=1\linewidth]{../results/synthetic_templates}
	\caption{Example of synthetic dechirped echo sequences, their corresponding templates and the templates reconstructed from the \pca\ \pcs. (xx-xx) Three examples of synthetic echo sequence waveforms. Panels xx-xx depict the resulting templates (i.e., summation across the output channels of the \citet{Wiegrebe2008} model). These panels also show the result of projecting the templates into a \pca D \pc\ space and transforming the result back to the original coordinates.}
	\label{fig:synthetictemplates}
\end{figure*}

In this section, we  used a generative model for naturalistic echoes. We assumed that echoes sequences received by bats can be considered as generated by anisotropically distributed point reflectors. Using this model, we have derived a low dimensional space ($D = \pca$), which can be used to describe the templates resulting from the artificial echo sequences. This implies that, after filtering through (a model of) the bat's auditory system, these artificial echoes can be converted into low dimensional descriptions. The original echo sequences, generated by xx to xx point reflectors effectively map to \pca\ dimensional vector. In the remainder of the paper, we will assess whether this low dimensional space allows describing the echoes from real bat habitats as collected using an ensonification device.

@@

In our previous work, the ensonification device was used to ensonify bat habitats at three different sites. However, here we only use data collected along two linear pathways through natural habitats. At the first site, a public park in Sde Boker, Israel (Fig. \ref{fig:datacollection}). The environment consisted of a walkway lined with rocks and vegetation. Unpublished bat monitoring data at the site confirmed that bats used this corridor to commute. The ensonification device was placed at 50 positions along a straight line spanning a total of 10 meters Locations were spaced 20 cm apart. At each location, echoes were collected for 31 azimuth directions (-150 to +150 degrees) and 7 elevation directions (-40 to +25 degrees). Therefore, at each location, echoes were collected from 31 x 7 = 217 directions. At each azimuth and elevation direction, three repeat measurements were recorded. In this way, a total of 32,550 echo trains (50 positions x 217 directions x 3 repeats) for each of the 31 microphones in the device were collected.

@@

As we did for the synthetic echo sequences, the echo trains collected in bat habitats were dechirped and filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}. We concluded from analyzing the synthetic echoes that the response in different frequency channels is highly correlated. Therefore, the output of the cochlear model was averaged across frequency channels to obtain a time-intensity profile, or template. As the beam of the emitter used was more narrow than the typically combined hearing and emission directionality of bats \citep{Vanderelst2010a,Reijniers2010,Jakobsen2012} templates were averaged across three neighboring directions. The first 6 ms of data were removed to avoid using the emissions themselves for analysis. The processing steps outlined above resulted in 3 $\times$ 217 templates, for each location at each site, corresponding to 217 directions and three repeats.

Among the set of templates, some had very little energy in them, notably those collected for directions without reflectors. For example, aiming the ensonification device upwards often resulted in no echoes returning. These templates could not be classified \citep{Vanderelst2016}. Therefore, we removed them from the current analyses by omitting all templates with very low energy. The resulting number of templates included in the present analyses are given in table \ref{tab:sizes}.

@@

Internal (and, potentially, external) noise results in template values greater than zero, even in the absence of reflectors returning echoes. To avoid training the neural network on this noise, we estimated the noise floor in our data by collecting templates while pointing the device to empty space. The maximum template value obtained from these template measurements was used as the noise floor $\noisefloor$. The value of the noise floor $\noisefloor$ used had a numerical value of 0.15. Any template values below this value were set to this value. As we argued before \citep{Vanderelst2016}, using the empirical measurement noise to estimate $\noisefloor$ is biologically warranted as bats have a better signal to noise ratio in their auditory system compared to our measuring device. Indeed, their dynamic range is at least 30 \db\ larger than our measurement device.

@@

We assessed whether the templates projected into the \pca-dimensional space still contained sufficient information to be recognized. Previously, we assessed whether templates can be used to recognize locations using, what \citet{Baddeley2012} called, a perfect memory approach. Templates were stored in a lookup table. The Euclidean distance between two echo templates was the metric used to assess how well bats should be able to distinguish between templates. This table lookup and distance calculation requires a comprehensive search of memory that scales with the number of templates memorized. Hence, in our previous work, we established the theoretical possibility of classifying templates based on real echoes from environments bats were known to frequent, irrespective of the significant amount of memory and computational resources required to do so.  Here, we use a more biologically plausible memory system: an artificial neural network, i.e. a neurally inspired way of distributed information storage \citep{Mcleod1998}. While we acknowledge that current neural network training methods most likely have no biological counterpart, we propose that this approach provides us with a more biologically plausible appraisal of the memory and computational effort required to implement a similar scheme in a bat brain.

Simple artificial feedforward neural networks were implemented using Tensorflow 2 for Python. The first layer of the neural network added Gaussian noise to each input sample ($\sigma(\stochasticnoise) = 0.01$). The following layers were fully connected and used the rectified linear activation function. The output layer uses a softmax activation function, i.e. the normalized exponential function. The loss was calculated as the cross-entropy between the target output and the actual output of the neural network. Weights were updated after each epoch using the Adam algorithm \citep{Kingma2014} .

@@

Training a different neural network for each of these features ensured that the training algorithm paid equal attention to each feature. When training a single network for all features, the optimization procedure would preferably reduce errors in the location predictions as this feature has the largest variance. The full architecture can be considered as a single network with different streams for each feature (fig. \ref{fig:networks}).

@@

We found that the templates derived from acoustic data collected in real habitats could be well represented by the \pca\ components derived from synthetic data. This \pca-dimensional space could represent xx\% of the variance in the templates. Figure \ref{fig:empiricaltemplates} shows examples of templates and their reconstructed counterparts. Therefore, although the model used to generate echo sequences was very simple, these results indicate that the \pca D feature space derived from the generative echo model could adequately represent templates resulting from real acoustic data.

@@

In our previous work \citep{Vanderelst2016}, we showed that the templates could be used to recognize locations in bat habitats. As we have shown that the templates can be projected into a lower \pca D feature space without loss of information, it can be expected that these \pca\ features can be used to recognize locations as well. We found this to be indeed the case.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\linewidth]{../results/royal_results.pdf}
	\caption{Classification results for the Royal Fort Park site. Top row: cumulative error distributions for the azimuth, elevation and location classification. Middle row: Confusion matrices for the classifications. Bottom row: remaining classification entropy for each target (derived from the confusion matrices).}
	\label{fig:royalperformance}
\end{figure*}

@@

\citet{Reijniers2010a} quantified the transmission of auditory information in the cochlea of bats. In that study, the authors adapted the Meddis inner hair cell model \citep{Meddis2006} to assess the amount of information in the acoustic signal preserved by the cochlea. The authors concluded that spectral notches introduced by two overlapping echoes are faithfully encoded by auditory nerve spike patterns. However, quantifying the information transfer rate, they also found that about half of the information (in terms of bits) are lost in the transduction from acoustic energy to spike trains. In the current paper, we expand on this work and quantify the dimension reduction due to cochlear encoding. Moreover, we show that this reduction in complexity still supports an ecologically valid task: sonar-based scene recognition.

We assumed echoes received by bats can be modeled as originating from a large collection of point reflectors. Passing these synthetic echoes through a model of the auditory periphery, we obtained a time-intensity profile (template) for each echo sequence. These templates could be well described by a low dimensional space of \pca\ principal components (figure \ref{fig:synthetictemplates}). We found that the same \pca D space could be used to describe templates originating from echoes collected in real bat habitats (figure \ref{fig:empiricaltemplates}). This suggests that the filtering of the auditory periphery provides the bat with a significantly compressed acoustic description of its environment. To identify the locus of the compression, we performed a number of additional analyses. In particular, we evaluated whether the low-dimensional description is a property of the echo sequences themselves, or whether this property results from the finite bandwidths of the gammatone filters and/or the low-pass filter in the \citet{Wiegrebe2008} model.

@@

Figure \ref{fig:mechanism} shows the number of \pcs\ needed to capture 99\% of the templates variance. From this plot, it can be seen that the number of \pcs\ increases when shuffling the echo sequences from xx to xx. Omitting the 1 kHz filter from the  \citet{Wiegrebe2008} model increases the number of \pcs\ to xx. Inspecting the average spectrum of the templates (fig. \ref{fig:mechanism}, inset) shows that most energy in the templates is contained in frequencies below 1 kHz. Therefore, if we take 1 kHz as the highest relevant frequency, it ought to be possible to completely represent templates by sampling them at 2 kSamples/sec. This corresponds to about 70 samples per template, as templates consist of 7499 samples at 219 kHz sample rate. This corresponds roughly to the number of \pcs\ required to represent templates derived from echo sequences that were shuffled  (fig. \ref{fig:mechanism}). The number of \pcs\ required to represent unshuffled templates is about half of this, i.c., \pca. Therefore, we conclude that the 1 kHz filter is the main, but not exclusive, locus of the dimensionality reduction. The structure of the echo signals also decreases the dimensionality of the templates. Therefore, we conclude that the sonar system of bats, through a number of filtering stages, compresses the 3D complexity of the world onto a lower dimensional representation.

@@

Despite the low dimensional character of the templates, they are sufficient to recognize environments. Indeed, this was already demonstrated in our previous work \citep{Vanderelst2016}. Here, we directly demonstrated that the \pca\ principal components are sufficient to discriminate the associated locations and viewing directions using neural networks. The neural networks' performance was somewhat that the perfect memory approach we used previously (figs. \ref{fig:israelperformance}, \ref{fig:royalperformance}).

While recognizing location in the environment is a critical component of navigation, bats use their sonar not only for navigating. They also rely on sonar for flight control. This includes, avoiding obstacles and/or following guiding structures. The environmental description the bat has access to for flight control starts from the same cochlear output  used for place recognition. Therefore, our current work suggests that the same low-dimensional description should be sufficient for flight control as well. Indeed, we have previously demonstrated that a low-dimensional description is sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}. In that work, we converted both the echoes at the left and at the right ear into a template-like representation used as input to the motion controller. Subsequent work will have to demonstrate whether the parameters of the \pca D space presented here can be used directly for controlling flight. In other words, we conjecture it should be possible to extract \pca\ features from an echo signal and convert this vector into a motor command. One could presume a neural network that converts the \pca\ features to flight control parameters.

@@

We were unable to find values for the number of neurons in the cortex of bat species. However, the mouse cortex has been estimated to contain about $9 \cdot 10^4$ neurons or $7 \cdot 10^8$ synapses per cubic millimeter \citep{Braitenberg2013}. The neural networks presented in this paper consisted of about \nneurons\ neurons and \nweights\ connections. Therefore, to the extent that the proposed neural networks can be considered as biologically plausible models of memory, they are orders of magnitude smaller than a single cubic millimeter of mouse cortex. Let alone the complete cortex, which has been estimated to contain 75 million neurons \citep{Williams2000}.

@@

It has been suggested before that bats' sonar systems provides detailed information about the world. A number of authors have referred to sonar as providing the bat with an acoustic image \citep{Lee2017,Barchi2013,Moss2001,Schnitzler2003,Simmons2012,Ulanovsky2008,Clare2015,Surlykke2016,Geipel2013}. As mentioned above, this is at odds with some recent behavioral and neurophysiological findings in bats faced with complex echo sequences \citep{Knowles2015,Geberl2013,Warnecke2018}. However, it is also at odds with the various filtering stages involved in the emission and the reception of bio-sonar. In particular, we argued before that the neurophysiology of the  bats auditory system represents a bottleneck in the amount of raw echo signal information accessible to the bat. Indeed, \citet{Reijniers2010a} already noticed that significant compression occurs during cochlear transduction. Here, we followed up on that work to quantify the dimensional reduction of the input signals as a result of peripheral auditory processing.

From our current work, in combination with previous results \citep{Vanderelst2016,Vanderelst2015a,Mansour2019,Reijniers2010a}, we conclude that the sonar system of bats maps complex echo signals onto a limited number of features. These features can be efficiently stored to recognize scenes. Moreover, we conjecture that a (binaural) version of the features should allow for sonar-based flight control. As such, a view of echolocating bats living in a low dimensional sensory space emerges that is at odds with the widespread notion of sonar as a sensing modality that provides detailed acoustic images.
